My deepest gratitude and thanks go to my advisor, Dr. Mukesh Dhamala, for all the exciting research opportunities, expert guidance, constant encouragement and emotional support throughout my PhD journey. The knowledge and experiences I gained working with Dr. Dhamala spans beyond this study. I am equally indebted to my co-advisor, Dr. Charles M. Epstein for his insightful research ideas, invaluable guidance and unwavering support that were instrumental in the success of this dissertation. 

Earnest thanks to my dissertation committee members, Dr. Thoms and Dr. Lei, for reviewing my work and providing valuable comments and suggestions. Special thanks to Dr. Bhim Adhikari for regularly attending our research meetings and providing research inputs. 

I greatly appreciate and acknowledge the support received from Brain and Behavior fellowship throughout my PhD degree. Sincere thanks to the faculty and staffs of the Department of Physics and Astronomy, my colleagues - Kiran and Tim, and my friends in the Physics and Astronomy department for their support in all possible ways.

Sincere gratitude to my parents (Keshav and Harimaya), my sister Radha, and brother Pramod for being a source of inspiration and strength. Thanks to Madhu babu and Sanju for supporting me in many different ways. I cannot express enough love and gratitude to my husband Madhav for his unconditional love, support and encouragement in all of my pursuits. To my dear son Sulav and daughter Isha, you make me fulfilled, bring enormous joy and meaning in my life. I love you to the moon and back.
